Supporting Information: Improving cropland maps through tight integration of human and machine intelligence

### 2020-08-14 inst/manuscript/figures.Rmd
Label risk assessment completed and integrated into SI and main text.

```{r, eval=FALSE}

res <- 0.005
res_supergrid <- 1
num_supergrid <- res_supergrid / res

@@

### Figure S4: Initial training site distribution
Figure showing distribution of initial 500 samples per cluster, as well as distribution of points per iteration.  
```{r trainingsites, eval=FALSE}
data("train_val_sites")

@@

### Figure S5: Label count and quality
```{r labelquality, eval=FALSE}
data("label_summary")

@@

### Figure S6: Consensus label (Bayesian) risk

[Not complete: checking values]
```{r, eval=FALSE}
data("label_consensus")
label_consensus_pts <- label_consensus %>% 
  rename(Risk = consensus_conflict) %>% 
  st_as_sf(coords = c("x", "y"), crs = 4326)
# label_consensus %>% distinct(aoi)

xylims <- st_coordinates(ghana)
maprange <-  c(range(xylims[, 1]), range(xylims[, 2]))
pal <- colorRampPalette(rev(brewer.pal(9, "RdBu")))

p <- ggplot(africa) + 
  geom_sf(fill = "gray97", color = "grey", size = 0.25) +
  geom_sf(data = aois, fill = "grey85", col = "black", size = 0.3) +
  geom_sf(data = label_consensus_pts, aes(color = Risk, fill = Risk), 
          size = 0.1) +
  scale_color_gradientn(colours = pal(20)) + 
  scale_fill_gradientn(colours = pal(20)) +
  coord_sf(xlim = maprange[1:2], ylim = maprange[3:4]) +
  guides(fill = guide_colorbar(nbin = 20), color = FALSE) +
  theme_void() + 
  theme(legend.key = element_rect(fill = "grey85", color = "transparent"))
p
label_consensus %>% rename(Risk = consensus_conflict) %>% 
  ggplot() + geom_histogram(aes(x = Risk)) + 
  facet_wrap(vars(aoi))

ggsave("manuscript/figures/si_map_reference_sample.png", width = 4, 
       height = 4, units = "in", dpi = 300, bg = "transparent")

>>>>

```{r, eval=FALSE}
res <- 0.005
res_supergrid <- 1
num_supergrid <- res_supergrid / res

@@

### Figure S3: Initial training site distribution
Figure showing distribution of initial 500 samples per cluster, as well as distribution of points per iteration.  
```{r trainingsites, eval=FALSE}
data("train_val_sites")

@@

### Figure S4: Label count and quality
```{r labelquality, eval=FALSE}
data("label_summary")

@@

### Figures S5 and S6: Consensus label (Bayesian) risk

Plots of mean risk in space, divided into training and validation sites

```{r riskpoints, eval=FALSE}
data("label_risk")

# Training/validation points
label_risk_pts <- label_risk$pts %>% 
  st_as_sf(coords = c("x", "y"), crs = 4326)
# label_consensus %>% distinct(aoi)
# label_risk %>% filter(is.na(meanrisk))

# # prepare for histograms
# clusters <- paste0("labeller", c("", "trainsouth", "trainsouthnl", "trainsw"))
# new_cluster_nms <- paste0("Cl", c("1", "2a", "2b", "3"))
# 
# label_risk_tb <- as_tibble(label_risk_pts) %>% 
#   mutate(aoi = case_when(
#     aoi == clusters[1] ~ "-3",
#     aoi == clusters[2] ~ "-2",
#     aoi == clusters[3] ~ "-1",
#     aoi == clusters[4] ~ "0", 
#     TRUE ~ gsub("labeller", "", aoi) 
#   )) %>% 
#   mutate(ord = as.numeric(aoi)) %>% arrange(aoi) %>% 
#   mutate(aoi = case_when(
#     aoi == "-3" ~ new_cluster_nms[1], 
#     aoi == "-2" ~ new_cluster_nms[2],
#     aoi == "-1" ~ new_cluster_nms[3],
#     aoi %in% "0" ~ new_cluster_nms[4], 
#     TRUE ~ aoi 
#   )) %>% 
#   mutate(aoi = forcats::fct_reorder(aoi, ord))
# 
# # summarize
# label_risk_stats <- label_risk_tb %>% group_by(aoi, usage) %>% 
#   summarize(Risk = mean(Risk, na.rm = TRUE)) %>% 
#   ungroup()

xylims <- st_coordinates(ghana)
maprange <-  c(range(xylims[, 1]), range(xylims[, 2]))
pal <- colorRampPalette(rev(RColorBrewer::brewer.pal(9, "RdBu")))
zlim <- c(0, round(max(label_risk_pts$Risk, na.rm = TRUE), 1))

p <- ggplot(africa) + 
  geom_sf(fill = "gray97", color = "grey", size = 0.25) +
  geom_sf(data = aois, fill = "grey85", col = "black", size = 0.3) +
  geom_sf(data = label_risk_pts, aes(color = Risk, fill = Risk), 
          size = 0.1) +
  facet_wrap(~usage) + 
  scale_color_gradientn(colours = pal(20), limits = zlim) + 
  scale_fill_gradientn(colours = pal(20), limits = zlim) +
  coord_sf(xlim = maprange[1:2], ylim = maprange[3:4]) +
  guides(fill = guide_colorbar(nbin = 20), color = FALSE) +
  theme_void() + 
  theme(legend.key = element_rect(fill = "grey85", color = "transparent"))
ggsave(pth("figures/si_label_risk_map.png"), width = 7, 
       height = 5, units = "in", dpi = 300, bg = "transparent")
```

Risk histograms. 
```{r riskhists, eval=FALSE}
p <- ggplot(label_risk$pts) + 
  geom_histogram(aes(x = Risk), fill = "grey50", bins = 20) + 
  geom_vline(data = label_risk$stats, aes(xintercept = Risk), size = 0.5, 
             linetype = 1) + 
  scale_y_continuous(expand = c(0, 0), breaks = seq(0, 150, 25), 
                     labels = c("", 25, "", 75, "", 125, ""), 
                     limits = c(0, 150)) + 
  ylab("Number of sites") + 
  facet_grid(aoi ~ usage) +
  theme_linedraw() + 
  theme(strip.background = element_rect(fill = "white"), 
        strip.text = element_text(size = 9, colour = "black"),  
        axis.text.x = element_text(size = 9, hjust = 1, vjust = 0.5, 
                                   angle = 90), 
        axis.text.y = element_text(size = 7))
ggsave(pth("figures/si_label_risk_hists.png"), width = 3, 
       height = 9, units = "in", dpi = 300, bg = "transparent")
```

### 2020-08-14 inst/manuscript/manuscript.Rmd
Label risk assessment completed and integrated into SI and main text.

Where $\theta$ represents the true cover type of a pixel (field or not field), *D* is the worker's label of that field, and $W_i$ is an individual worker.  P($\theta$|D) is therefore the probability that the actual cover type is what the workers who labelled it says it is, while P(W$_i$|D) is the worker's average score over the accuracy assessment assignments they have mapped to date, and P(W$\theta$|D, *W$_i$*) is the worker's label for that pixel. This approach therefore uses the overall accuracy of each worker to weight their labels as they are combined with other workers' labels for a given pixel (see SI for further details). **The "risk" of the final consensus label is then assessed, which measures the proportion of pixels in the label exceeding a Bayesian risk threshold (see SI), and thus provides an additional measure of label quality.**

@@

After each site was mapped by four labellers, consensus labels were generated. **The average risk for each consensus label was *X* (see SI)**.

To further evaluate label quality, we assessed the degree of similarity between worker polygons with the field boundaries collected by our ground validation team. Overall, we found that on average a worker's field boundary interpreted from a PlanetScope image had only an **X%** agreement (as determined by area of intersection over union) with the corresponding field-mapped boundary, and the average field size mapped by a worker was on average **X% smaller/larger** than those mapped on the ground.

>>>>

Where $\theta$ represents the true cover type of a pixel (field or not field), *D* is the worker's label of that field, and $W_i$ is an individual worker.  P($\theta$|D) is therefore the probability that the actual cover type is what the workers who labelled it says it is, while P(W$_i$|D) is the worker's average score over the accuracy assessment assignments they have mapped to date, and P(W$\theta$|D, *W$_i$*) is the worker's label for that pixel. This approach therefore uses the overall accuracy of each worker to weight their labels as they are combined with other workers' labels for a given pixel (see SI for further details). As a further measure of confidence in the final consensus label, its average Bayes Risk is calculated (see SI). This measure ranges between 0 and 1, with 0 indicating full agreement between labellers for all pixels (n = 40000) in the label, and 1 indicating complete disagreement.

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
data("label_risk")
murisk <- label_risk$stats %>% group_by(usage) %>% 
  summarize(Risk = round(mean(Risk), 3)) %>% 
  tidyr::pivot_wider(names_from = usage, values_from = Risk)
mintrainrisk <- label_risk$stats %>% filter(usage == "Train") %>% 
  filter(Risk == min(Risk))
maxtrainrisk <- label_risk$stats %>% filter(usage == "Train") %>% 
  filter(Risk == max(Risk))
minvalrisk <- label_risk$stats %>% filter(usage == "Validate") %>% 
  filter(Risk == min(Risk))
maxvalrisk <- label_risk$stats %>% filter(usage == "Validate") %>% 
  filter(Risk == max(Risk))

# label clusters
cl1 <- label_risk$stats %>% filter(aoi %in% 1:6) %>% group_by(usage) %>% 
  summarize(Risk = round(mean(Risk), 3))
cl2 <- label_risk$stats %>% filter(aoi %in% c(7:9, 12, 15)) %>% 
  group_by(usage) %>% summarize(Risk = round(mean(Risk), 3))
cl3 <- label_risk$stats %>% filter(aoi %in% c(10:11, 13:14, 16)) %>% 
  group_by(usage) %>% summarize(Risk = round(mean(Risk), 3))

# training clusters
cl2arisk <- label_risk$stats %>% filter(aoi %in% c("Cl2a")) %>% pull(Risk) %>% 
  round(3)
cl2brisk <- label_risk$stats %>% filter(aoi %in% c("Cl2b")) %>% pull(Risk) %>% 
  round(3)

After each site was mapped by four labellers, consensus labels were generated. The Bayesian Risk (see SI) of each training and validation label was calculated as an additional measure of label quality. The average risk was `r murisk$Train` for training labels and `r murisk$Validate` for validation labels. Risk was highest in the northern cluster (AOIs 1-6; Figure S5), falling between `r cl1$Risk[1]` for training and `r cl1$Risk[2]` for validation labels (Figure S6), and lowest in the southwestern cluster (AOIs 10, 11, 13, 14, 16; training risk = `r cl3$Risk[1]`; validation risk = `r cl3$Risk[2]`). Label risk in the central/eastern cluster (AOIs 7-9, 12, 15) was slightly lower (training = `r cl2$Risk[1]`; validation = `r cl2$Risk[2]`) than in the north. Greater labeller experience also appeared to reduce risk, as the mean risk for the second mapping of Cluster 2's 500 initial training sites was `r cl2brisk`, compared to `r cl2arisk` for the first mapping.

**[To further evaluate label quality, we assessed the degree of similarity between worker polygons with the field boundaries collected by our ground validation team. Overall, we found that on average a worker's field boundary interpreted from a PlanetScope image had only an *X%* agreement (as determined by area of intersection over union) with the corresponding field-mapped boundary, and the average field size mapped by a worker was on average *X% smaller/larger* than those mapped on the ground.]**

### 2020-08-14 inst/manuscript/si.Rmd
Label risk assessment completed and integrated into SI and main text.

documentclass: article
classoption: a4paper
capsize: normalsize
fontsize: 11pt
geometry: margin=1in
linenumbers: false
spacing: doublespacing
footerdate: yes
# abstract: The abstract can go either here or below
keywords: rmarkdown, reproducible science
bibliography: 
  - references.bib
  - knitcitations.bib
csl: ecology.csl

@@

After creating the consensus label, the degree of confidence in the resulting label value can be expressed in terms of Bayesian risk. Risk is calculated as:

@@

Where *C* is the consensus probability that a given pixel is a field ($\mathrm{P(\theta = field|D)}$), and *L* is the consensus label (i.e. non-field if *C* < 0.5, field if *C* > 0.5) for that pixel. The risk values across the entire sample site can be processed in two ways to provide useful information about the confidence in the consensus label for that site. The first is a simple average of all risk values in the site, where the slope of the risk varies depending on whether *L* is a field or not a field (Figure \ref{fig:riskcurve}). The closer to 0 the lower the risk that the *L* is mislabelled, while values approaching 1 indicate increasing risk of mislabelling. A second approach is to calculate the proportion of pixels having risk values that exceed a certain threshold. This second approach is currently implemented in our system, using a threshold value of 0.4.

Where *C* is the consensus probability that a given pixel *j* is a field ($\mathrm{P(\theta = field|D)}$), and *L* is the consensus label (i.e. non-field if *C* < 0.5, field if *C* > 0.5) for that pixel. *R* is thus the average risk across all pixels at a given site, and the slope of the risk varies depending on whether *L* is a field or not a field (Figure \ref{fig:riskcurve}). The closer to 0 the lower the risk that the *L* is mislabelled, while values approaching 1 indicate increasing risk of mislabelling.

@@

>>>>

documentclass: article
classoption: a4paper
capsize: normalsize
fontsize: 11pt
geometry: margin=1in
linenumbers: false
spacing: doublespacing
# abstract: The abstract can go either here or below
keywords: rmarkdown, reproducible science
bibliography: 
  - references.bib
  - knitcitations.bib
csl: ecology.csl

@@

After creating the consensus label, the degree of confidence in the resulting label value is measured by Bayes Risk:

@@

Where *C* is the consensus probability that a given pixel is a field ($\mathrm{P(\theta = field|D)}$), and *L* is the consensus label (i.e. non-field if *C* < 0.5, field if *C* > 0.5) for that pixel. The risk values across the entire sample site can be processed in two ways to provide useful information about the confidence in the consensus label for that site. The first is a simple average of all risk values in the site, where the slope of the risk varies depending on whether *L* is a field or not a field (Figure \ref{fig:riskcurve}). The closer to 0 the lower the risk that the *L* is mislabelled, while values approaching 1 indicate increasing risk of mislabelling. A second approach is to calculate the proportion of pixels having risk values that exceed a certain threshold.

@@

The average Bayes Risk of each training and validation site is shown in Figure \ref{fig:labelrisk}, as well as the distribution of risk values per AOI and the initial training clusters (Figure \ref{fig:labelriskhist}). The three initial clusters include the second mapping of Cluster 2.

```{r labelrisk, echo = FALSE, out.width="80%", fig.cap="The average Bayes Risk of each training and validation site in Ghana.", fig.align='center', message=FALSE, results='hold'}
knitr::include_graphics('figures/si_label_risk_map.png')
```

```{r labelriskhist, echo = FALSE, out.width="45%", fig.cap="The distribution of Bayes Risk values for training (left column) and validation (right column) sites in each AOI, with the average value indicated by vertical lines. The first three rows (panel titles beginning with 'Cl') indicate distributions of Bayes Risk in the initial training clusters, including the first ('Cl2a') and second ('Cl2b') mappings of Cluster 2.", fig.align='center', message=FALSE, results='hold'}
knitr::include_graphics('figures/si_label_risk_hists.png')
```
